### 2015-05-09 paper/abstract1.tex
Work on the abstract

\abstract{
    To achieve consistent performance, a distributed computing service must
    dynamically adapt to unpredictable crashes, communication failures, and
    changes in the availability of shared resources.
    Unfortunately, most existing algorithms are not adaptive: outside of their
    sweet spot, they have poor performance.

    We present a new abstraction, Abortable Generalized Consensus, which enables the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms.

    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms. We provide two specifications satisfying
    ACA and we illustrate their power by refining them to
    obtain a versatile adaptive algorithm that combines five existing
    algorithms into a single one. ACA is completely formalized and its
    underlying composition theorem has been mechanically checked by an interactive
    proof assistant.
}