### 2015-05-14 paper/roadmap.tex
Moved the intro and roadmap to their own files.

The rest of the paper is structured as follows.
\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
presents Abortable Chain Agreement and its optimizations to treat commuting
and read-only commands efficiently. \Cref{sec:rounds} presents our specification
of conservative and fast ACA rounds, which implement ACA. In
\cref{sec:examples}, we show that the rounds of  
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos can easily be transformed into ACA implementations
by using our specifications of conservative and fast ACA rounds.
Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.