### 2016-01-06 paper/abstract3.tex
Switched to the llncs template. Reworked the abstract and intro.

\begin{abstract}
State-Machine Replication (SMR) is a basic building block of reliable
distributed systems. Improving SMR algorithms used by cloud-computing providers
can have a big impact on the reliability of the many computing services they
support. However, doing so is hard, often justifying a publication in a top
conference.

The problem is that SMR algorithms are not modular. Thus one cannot optimize an
existing algorithm for a specific situation without having to redo from scratch
all the analysis, proofs, and tests that were done for the existing algorithm.

We present a framework that enables building SMR algorithms out of independent
modules. A module is an SMR algorithm that can abort its execution and pass
the baton to a new module, allowing to build composite SMR algorithms which
can dynamically switch between modules. Crucially, each module can be analyzed
once and for all in isolation; our composition theorem guarantees that the
composition of any number of correct modules is a correct SMR algorithm.

The key novelty of our framework is its wide applicability: we show that a
large class of SMR algorithms can be expressed in our framework by proposing
two specifications capturing the essence of modular SMR algorithms based on
quorum systems. We demonstrate that concrete algorithms can easily be obtained
by refinement from our abstract specifications and we show that our framework
enables formal specification and verification of composite SMR algorithms in
TLA+ and Isabelle/HOL\@.
\end{abstract}