### 2016-01-06 paper/abstract4.tex

\begin{abstract}

State-Machine Replication (SMR) algorithms such as Paxos are notoriously hard to
understand, let alone to devise. Moreover, there is no practical way to reuse and optimize existing SMR
algorithms, therefore the cost of creating a new SMR algorithm is high.

To lower this cost, we present a framework for building SMR algorithms out of independent and reusable SMR modules. A module is an SMR
algorithm that can abort and pass the baton to another module, allowing to dynamically switch between modules. 
We propose Composable
Generalized Consensus (CGC), a correctness condition for individual modules, and
we prove that the composition of correct modules is a correct SMR algorithm.
Our framework consists of TLA$^+$ specifications of CGC and of two abstract
types of correct modules, fast rounds and conservative rounds, each providing a template to devise new SMR modules with the help of the TLA$^+$
tools. We show that several well-known SMR algorithms can be modified to
implement fast or conservative rounds, therefore enabling
their composition to obtain a single versatile SMR algorithm.

%With our specifications, algorithm designers can use the TLA$^+$ tools
%to prototype and debug new SMR algorithms. Moreover, we also provide a
%specification of CGC and a proof of the composition theorem in Isabelle/HOL\@.

\end{abstract}