### 2016-01-07 paper/abstract4.tex
Work on the abstract.

\begin{abstract}

State-Machine Replication (SMR) algorithms such as Paxos are notoriously hard to
understand, let alone to devise. Moreover, there is no practical way to reuse
and optimize existing SMR algorithms, therefore the cost of creating a new SMR
algorithm is high.

To lower this cost, we present a framework for building SMR algorithms out of
independent and reusable SMR modules. A module is an SMR algorithm that can
abort and pass the baton to another module, allowing to dynamically switch
between modules. Our composition theorem ensures that the composition of any
number of modules satisfying Composable Generalized Consensus (CGC), our
correctness condition for modules, is a correct SMR algorithm.

To help design new modules in practice, we propose TLA$^+$ specifications
of CGC and of two abstract types of modules, fast rounds and conservative
rounds, each providing a template to devise new SMR modules with the help of the
TLA$^+$ tools. We show that several well-known SMR algorithms can be modified to
implement fast or conservative rounds, therefore enabling their composition to
obtain a single versatile SMR algorithm.

%With our specifications, algorithm designers can use the TLA$^+$ tools
%to prototype and debug new SMR algorithms. Moreover, we also provide a
%specification of CGC and a proof of the composition theorem in Isabelle/HOL\@.

\end{abstract}