### 2016-01-14 paper/intro.tex

\section{Introduction}

The State-Machine Replication technique (abbreviated SMR) allows building
reliable computing services on top of unreliable hardware. The basic idea
is to replicate a service over several replica servers and enforce, using
a distributed algorithm, that each replica executes the same sequence of
deterministic commands. In this fashion, a command will produce the same effect
on all replicas. Therefore, if a replica crashes, clients of the service can
be transparently redirected to another replica and will have the illusion
of accessing a centralized service at all times. Formally, we will use the
Generalized Consensus specification~\cite{Lamport05GeneralizeConsensus},
abbreviated GC, as our correctness condition for SMR algorithms.

SMR is part of the basic infrastructure supporting cloud computing; therefore,
improving the performance of SMR algorithms can result in performance gains,
cost reduction, or improved resilience (because SMR may be used where it was
too costly before) for many online services.

However, devising new SMR algorithms is very costly because one has to employ
formal verification techniques to ensure their correctness. SMR algorithms are
notoriously difficult to understand and it is easy to overlook catastrophic bugs
appearing not only in implementations but also in the high level algorithm.
Moreover, concurrency, network behavior, and faults give rise to a number
of possible interleavings of actions that is beyond the reach of analysis
by testing method. Therefore, to gain confidence in the correctness of an
SMR algorithm, one needs to resort to formal methods like model-checking or
interactive theorem proving. However, this is costly: model-checking often
requires manually building an abstraction of the algorithm to simplify the task
of the model checker, and interactive theorem proving is very time consuming.
Both require experts trained in the particular tool used for the task. As
evidence of the difficulty of designing new SMR algorithms, note that today,
more than two decades after the first SMR algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} 
were proposed, devising a new SMR algorithm can justify a publication in a top
conference~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

Surprisingly, at a first glance, many SMR algorithms seem to share a common
structure which would indicate that previous formal analysis could be reused.
This similarity in structure exists because, in many cases, researchers obtained
a new SMR algorithm by optimizing an existing SMR algorithm for a particular use
case. For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds
to MultiPaxos~\cite{lamport2001paxos} to reduce the latency under low
contention, Disk Paxos~\cite{GafniLamport03DiskPaxos} optimizes Paxos
for systems where disk nodes and processor nodes are distinct, Chain
Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability} 
organizes the replicas of Paxos in a ring to distribute
commands at a higher throughput but at the expense of latency,
FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus} optimizes
Generalized Paxos\cite{Lamport05GeneralizeConsensus} to reduce latency
when the system contains a reliable and fast leader node, Ring
Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol}
optimizes the latency and throughput of Paxos when network multicast is
available, Multicoordinated Paxos reduces the performance impact of a
leader fault in MultiPaxos by implementing a distributed leader, Cheap
Paxos~\cite{LamportMassa04CheapPaxos} optimizes Paxos to decrease the load
on some replicas in case failures are sufficiently separated in time,
Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine} 
optimizes MultiPaxos for wide-area networks by rotating the role of leader
among all replicas, etc. Close scrutiny reveals that most SMR algorithms can
be seen as variations and optimizations around a few core ideas such as the
round structure introduced by Paxos~\cite{Lamport98ParttimeParliament}, quorum
systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos},
and agreement up to commutativity based on command-structures~\cite{Lamport05GeneralizeConsensus}.

Given the similarities shared by many SMR algorithms, one may ask why is it
not possible to reuse formal analyses to simplify the verification of new
algorithms. The problem is that SMR algorithms are not modular and that, because
of their intricate structure, modifying a monolithic SMR algorithm may produce
many unforeseen corner cases that invalidate previous analyses.

To simplify the task of devising new SMR algorithms we propose a framework for
building self-contained and reusable SMR modules whose analysis is independent
of the other modules they may be composed with. For example, in \cref{}, we
augment MultiPaxos with our composition interface, and we prove that it satisfies
our correctness condition for modules, called Composable Generalized Consensus (CGC).
We then build an SMR module whose structure is similar to the one of the fast
rounds of Fast Paxos, and we prove that it satisfies CGC\@. Finally, we assemble
unmodified our MultiPaxos SMR module and our Fast Paxos SMR module in an SMR
algorithm that dynamically switches from one to the other depending on the
contention level observed in the system. This composite algorithm combines the
advantages of both its SMR modules: in cases of low contention, our Fast Paxos
module executes commands in two message delays, while MultiPaxos would need
three, whereas the MultiPaxos module maintains its throughput in high contention
cases, in which the performance of the Fast Paxos module would have been
significantly reduced.

Crucially, we obtain the correctness of our composite
SMR algorithm for free: our generic composition theorem and the correctness
of our two SMR modules taken individually imply that the combination of the
MultiPaxos SMR module and the Fast Paxos SMR module is a correct SMR algorithm.
Later, in \cref{}, we present three other SMR modules that can be combined the
MultiPaxos SMR module and the Fast Paxos SMR module in a composite SMR algorithm
that can switch between 4 different kinds of SMR modules, according to the
system's operating conditions.

More generally, an SMR module is an SMR algorithm augmented with a \emph{composition interface}
and satisfying a correctness condition that we call \emph{Composable Generalized
Consensus}, abbreviated CGC\@. The composition interface consists of an \emph{abort interface}
for aborting an SMR module's execution and passing the baton to another
SMR module and with a matching \emph{init interface} for taking over the
execution of an aborting SMR module. Two SMR modules are composed by passing
the outputs produced by the abort interface of the first SMR module, called
\emph{abort c-structs}, as inputs to the init interface of the second SMR module,
called \emph{init c-structs} in the context of the second SMR module. Any correct SMR module is a safe Generalized
Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR
algorithm), but the progress guarantees may vary between SMR modules: CGC only
requires an SMR module to make progress executing commands or to abort its
execution. CGC also relates the init c-structs, learned values (learned c-structs 
represent the local state-machine state of individual replicas), and abort
c-structs in order to guarantee that an SMR module consistently takes over the execution of its predecessor, as expressed in our composition theorem.
Finally, our framework allows building \emph{dynamic SMR
algorithms}, which dynamically change module when the currently running module
aborts, choosing the module best suited to the current operating conditions.

Our composition theorem enables incremental development of SMR algorithms
because it guarantees that any number of correct modules may be composed in a
sequence to form a correct SMR algorithm. Thanks to the composition theorem,
new modules can be added to a dynamic SMR algorithm without modifying the
existing modules and with guaranteed correctness. Moreover, the relaxed progress
requirements of CGC enable a separation of concerns that simplifies optimizing
SMR algorithms for particular scenarios because there is no need to consider
executing commands under a situation outside the scenario: the module can just
abort in this case.

In practice, we provide a TLA+ specification of CGC that the user of our framework can use to debug and prove new modules correct using the TLA$^+$ toolbox~\cite{TLATools}.
Moreover, we also provide two TLA+ specifications simplifying the development of modules based on Refined Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems}. These two specifications both implement CGC and provide a skeleton for a user to build her own module. The two specifications demonstrate how to update the local state of a replica after gathering the states or the init c-structs of a write quorum of replicas, and how to
produce an abort c-struct after gathering the state of a read quorum of replicas. Each of the specifications achieves a particular consistency/latency trade-off~\cite{Abadi12ConsistencyTradeoffsModernDistributedDatabaseSystem}.
The two specifications use different combinations of sizes of read and write quorums, which affects the invariants that the replicas must maintain and, in consequence, the latency with which a command can be processed but the module.
Our two specifications abstract over how to disseminate the commands to the replicas (e.g.\ whether to use a leader as in Paxos, multiple leaders as in MultiCoordinatedPaxos, or whether to broadcast commands without using a leader as in Fast Paxos).
The user can therefore specify a concrete command dissemination procedure and then model-check that the obtained algorithm refines one of the two specifications.  
We demonstrate the practical use of our framework by presenting modules based on Paxos~\cite{Lamport98ParttimeParliament}, Fast Paxos\cite{Lamport06FastPaxos}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, a case of Vertical Paxos~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.

Compared to related work, described in \cref{sec:related}, our key contributions are the following.
\begin{itemize}
    \item We present the first framework allowing to build Generalized Consensus algorithms our of independent and reusable modules supporting modular formal analysis.
    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing four CGC modules inspired by MultiPaxos, Fast Paxos, Chain Replication, and Primary-Backup replication.
\end{itemize}
These two contributions set this paper apart from related work, notably the Abstract framework~\cite{GuerraouiETAL10Next700BftProtocols} and the Speculative Linearizability framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}, both described in \cref{sec:related}.
Our TLA+ specifications can be found at \url{http://losa.fr/cgc}.
Moreover, we have also formalized the CGC correctness condition and proved the composition theorem in Isabelle/HOL\@, and the Isabelle theories are available at the same URL\@.